---
title: 'Appendix: Pilot Experiment'
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
require(lme4)
require(tidyverse)
require(here)
require(kableExtra)
require(sjPlot)
require(papaja)
require(lmerTest)

# for Bayesian regression
require(brms)
require(ggmcmc)
require(tidybayes)
require(ggdist)
require(modelr)
```

We add a new pilot dataset of 16 infants from Leiden University.
Notes: 
Julien's data has age rounded to month -- don't we want continuous age?
There are 16 data files for Leiden participants, but inside the file for participant 16 the subject ID is '15': is this a duplicate, or should just be changed to 16? 
What are the procedures used by each lab? (Leiden = HPP, Julien = eyetracking?)
Which random effects structures should we use for the 2 pilot labs (esp. tricky if they also vary procedure)?

## Participants

```{r load-julien-data}
d1 <- read_csv(here("pilot/data/first_session_babies.csv")) %>%
  rename(sd_LT_incongruent_trials = sd_LT_congruent_trials_1) 
d2 <- read_csv(here("pilot/data/second_session_babies.csv")) %>%
  rename(sd_LT_incongruent_trials = sd_LT_congruent_trials_1,
         looking_time = `looking time`) 

cdat <- bind_rows(d1, d2) %>%
  filter(!is.na(id)) %>%
  mutate(labID = 1,
         procedure = "eyetracking")
```

```{r process-leiden-data, eval=F}
demo <- read_csv("data/Data_Pilot_Leiden_Participants.csv") %>%
  select(1:26) # up to "notes" column

dem_cln = demo %>% mutate(age_months = `Age(days)` / 30.44) %>%
  rename(id = Subject, fam_condition=Condition, gender=Gender) %>%
  select(id, fam_condition, age_months, gender) %>%
  mutate(gender = ifelse(gender=="M", "male", "female"),
         labID = 2,
         procedure = "HPP") # ?
# also need to standardize language (L1, L2, etc) and amounts (L1H etc)


# varying numbers of columns per datafile
ldat <- tibble()

data_files <- list.files(path="data/Data_Pilot_Leiden", pattern='*.csv', full.names=T) 

for(fname in data_files) {
  tmp <- read_csv(fname) 
  # proper header
  names(tmp) = tmp[3,]
  tmp <- tmp[4:nrow(tmp),]
  
  # README: The relevant columns for looking times during test are:
  # `LookTime2.Block.`,`LookTime4.Block.`,  `LookTime6.Block.`, etc (only even numbers). 
  # The number of total look times depends on the number of separate looks that infants made.
  lt_cols <- which(startsWith(names(tmp), "LookTime"))
  lt_block_cols <- which(endsWith(names(tmp)[lt_cols], "[Block]"))
  test_indices = seq(2, length(lt_block_cols), 2)
  lt_col_names <- names(tmp)[lt_cols][lt_block_cols][test_indices]
  
  tmp_proc <- tmp %>% filter(AttenList.Cycle!="1") %>%
    select(ExperimentName, Subject, DataFile.Basename, Block, `AudioFile[Block]`, 
                 RandomSeed, lt_col_names) %>%
    mutate(across(where(is.character), ~na_if(., "NULL")))
  
  tmp_final <- tmp_proc %>% 
    mutate_at(lt_col_names, as.numeric) %>%
    mutate(looking_time = rowSums(.[grep("LookTime", names(.))], na.rm=T)) %>%
    select(-lt_col_names)
  
  ldat <- bind_rows(ldat, tmp_final)
}

# participant 16 was mislabeled as 15 (in the file MB3_ABB_16-1.csv the subject number is 15)
# fix based on RandomSeed (see email from Andreea)
ldat[which(ldat$RandomSeed=="-1749405221"),]$Subject = "16" 

ldat <- ldat %>% select(-RandomSeed) %>%
  rename(audio_file = `AudioFile[Block]`,
         trial = Block,
         id = Subject) %>%
  mutate(id = as.numeric(id)) %>% left_join(dem_cln)

save(ldat, file="data/merged_Leiden_pilot_data.Rdata")
```


```{r load-leiden-data}
load("data/merged_Leiden_pilot_data.Rdata")
```


```{r merge-data}

dat <- cdat %>% select(1:14, "labID") # remaining columns are averages and notes

fam_dat <- dat %>% filter(is.na(test_trial)) %>% 
  select(-test_trial, -test_trial_audio, -test_trial_audio_length_ms, 
         -test_trial_type, -test_version, -trial)

# Leiden familiarization data
fam_datl <- ldat %>% filter(str_detect(audio_file, "fam_")) %>% 
  rename(fam_trial_audio = audio_file) %>%
  mutate(id = as.character(id)) %>%
  select(-trial)

fam_dat_all <- fam_dat %>% select(intersect(names(fam_dat), names(fam_datl))) %>%
  bind_rows(fam_datl %>% select(intersect(names(fam_dat), names(fam_datl))))
# table(fam_dat_all$fam_condition) # 17 per condition

# Julien data
test_dat <- dat %>% filter(!is.na(test_trial)) %>% 
  select(-fam_trial_audio, -fam_version) %>%
  separate(test_trial_type, sep='_', 
           into = c("test_trial_type_ch", "test_trial_type_num")) %>%
  mutate(trial_type = ifelse(test_trial_type_ch == fam_condition, "same", "different"),
         trial = trial - 1) # was 2-13; now 1-12

# Leiden data
test_datl <- ldat %>% filter(!str_detect(audio_file, "fam_")) %>% 
  select(-DataFile.Basename, -ExperimentName) %>%
  rename(test_trial_audio = audio_file) %>%
  mutate(test_trial_type_ch = case_when(str_detect(test_trial_audio, "tisoti") ~ "ABA",
                                        str_detect(test_trial_audio, "tisoso") ~ "ABB",
                                        str_detect(test_trial_audio, "lonini") ~ "ABB",
                                        str_detect(test_trial_audio, "lonilo") ~ "ABA",
                                        str_detect(test_trial_audio, "lonini") ~ "ABB",
                                        TRUE ~ as.character(NA))) %>%
  mutate(id = as.character(id),
         trial_type = ifelse(test_trial_type_ch == fam_condition, "same", "different"),
         trial = as.numeric(trial) - 2) # was 3-14; now 1-12

needed_cols <- c("labID", "id", "gender", "age_months", "fam_condition", "trial_type", "trial", "looking_time")

test_dat_all <- test_dat %>% select(needed_cols) %>% 
  bind_rows(test_datl %>% select(needed_cols)) %>%
  mutate(log_looking_time = log(looking_time),
         age_sc = scale(age_months, center=T, scale=T)[,1],
         trial_sc = scale(trial, center=T, scale=T)[,1])
```

# Analysis

```{r, lmer-regression}
m1 <- lmer(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (trial * trial_type | id) + (1 | labID), 
           data=test_dat_all)
# isSingular(m1)
# singular, random effects need pruning (changing to trial_sc / age_sc did not help)

# try pruning either trial number or trial type (both converge)
m2 <- lmer(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (trial | id) + (1 | labID), 
           data=test_dat_all)
# what random effects structure with labID do we want? (+ (id | labID) is slow..)

m3 <- lmer(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (trial_type | id) + (1 | labID), 
           data=test_dat_all)

# anova(m2, m3) # m2 is slightly preferred

mod_out <- summary(m2)

papaja::apa_table(mod_out$coefficients, 
                  caption="Regression coefficients.")
```

```{r, fig.width=4.8, fig.height=3.5, fig.cap="Regression coefficients with 95% confidence intervals."}
plot_model(m2, sort.est = T) + theme_bw() + ggtitle("")  
#  geom_vline(aes(xintercept=0), linetype="dashed")
```

```{r log-looking-time-by-age, fig.width=4.5, fig.height=4, fig.cap="Log(looking time) by trial type and age, and bootstrapped 95% confidence intervals."}
dag <- test_dat_all %>% 
  group_by(id, fam_condition, age_months, trial_type) %>%
  #summarise(looking_time = mean(log_looking_time)) %>% 
  #group_by(trial_type, age_months) %>%
  tidyboot::tidyboot_mean(log_looking_time) # quite slow..

#pos = position_dodge(width=.2)
ggplot(dag, aes(x=age_months, y=mean, group=trial_type, color=trial_type)) + 
  geom_point(aes(y=mean, x=age_months)) + # , pos=pos
  ylab("log(looking time)") + xlab("Age (months)") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), alpha=.3) + #  pos=pos,
  theme_bw() + geom_smooth(method="lm") +
  scale_color_discrete(name="Trial Type")
```


## Bayesian Regression 

```{r, message=F, warning=F, eval=T, include=T}
bm0 <- brm(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             (1 | id) + (1 | labID), 
           data = test_dat_all, 
           warmup = 1000, iter = 2000, chains = 4, 
           inits = "random", cores = 4, seed=42) 

summary(bm0)

# add trial number effects
bm1 <- brm(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (trial | id) + (1 | labID), 
           data = test_dat_all, 
           warmup = 1000, iter = 2000, chains = 4, 
           inits = "random", cores = 4, seed=42) 

summary(bm1)

# Watanabe-Akaike Information Criterion
waic0 <- waic(bm0)
waic1 <- waic(bm1)
compare_ic(waic0, waic1)
# bm1 only a bit better than bm0 (2.75/5.39 = ~.5 SE lower)

```

```{r, echo=F, eval=T, include=T}
get_variables(bm1)
bdat <- ggs(bm1)

# bm1 %>% 
#   spread_draws(b_Intercept, b_fam_conditionABB, b_trial_typesame, b_age_months, b_trial,
#               `b_fam_conditionABB:trial_typesame`, `b_trial_typesame:age_months`,
#               `b_trial_typesame:trial`) %>%
#   mutate(mean = b_Intercept + b_fam_conditionABB * b_trial_typesame +
#            b_age_months * b_trial_typesame + 
#            b_trial * b_trial_typesame) %>%
#   ggplot(aes(y = fam_condition, x = mean)) +
#   stat_halfeye()

# Region of Practical Equivalence (ROPE)
bm1 %>%
  gather_draws(b_fam_conditionABB, b_trial_typesame, b_age_months, b_trial, # b_Intercept,
               `b_fam_conditionABB:trial_typesame`, `b_trial_typesame:age_months`,
               `b_trial_typesame:trial`) %>%
  ggplot(aes(y = .variable, x = .value, fill = stat(abs(x) < .8))) +
  stat_halfeye() +
  geom_vline(xintercept = c(-.8, .8), linetype = "dashed") +
  scale_fill_manual(values = c("gray80", "skyblue")) + 
  theme_classic()
  
# test_dat_all %>%
#   group_by(trial_type, id, labID, fam_condition, age_months) %>%
#   data_grid(age_months = seq_range(age_months, n = 101)) %>%
#   add_predicted_draws(bm0) %>%
#   ggplot(aes(x = age_months, y = log_looking_time, color = ordered(fam_condition), fill = ordered(fam_condition))) +
#   stat_lineribbon(aes(y = .prediction), .width = c(.95, .80, .50), alpha = 1/4) +
#   geom_point(data = mtcars) +
#   scale_fill_brewer(palette = "Set2") +
#   scale_color_brewer(palette = "Dark2")

```


Generate posterior predictive values. 

```{r, include=T, eval=T}
# posterior predictive
test_dat_all %>%
  data_grid(labID, id, fam_condition, trial_type, age_months, trial) %>%
  add_fitted_draws(bm1) %>%
  ggplot(aes(x = .value, y = trial_type)) +
  facet_grid(. ~ fam_condition) +
  stat_pointinterval(.width = c(.66, .95)) + 
  theme_bw()
```


